### Exponential smoothing with no trend

Simple exponential smoothing is like the naive model that just uses the last value to make the forecast, but instead of only using the last value it will use values farther in the past also. The weighting fall off exponentially.  The objective when fitting a simple exponential smoothing model is to estimate the best weighting ( $\alpha$ ):

```{r echo=FALSE}
alpha <- 1
wts <- alpha*(1-alpha)^(0:23)
plot(1987:1964, wts/sum(wts), lwd=2, ylab="weight", xlab="", type="l")
alpha <- 0.5
wts <- alpha*(1-alpha)^(0:23)
lines(1987:1964, wts/sum(wts), lwd=2, col="blue")
alpha <- 0.05
wts <- alpha*(1-alpha)^(0:23)
lines(1987:1964, wts/sum(wts), lwd=2, col="red")
legend("topleft", c("alpha=1 like naive","alpha=0.5","alpha=0.05 like mean"),lwd=2, col=c("black","blue","red"))
title("more weight put on more recent values\nfor 1988 forecast")
```

---

### Fit a ETS model

The `ets()` function in the forecast package fits a simple exponential smoothing model.

Load the data.

```{r load_data_exp_smoothing}
load("landings.RData")
```

Fit the model.

```{r fit.ann}
fit <- forecast::ets(anchovy87ts, model="ANN")
fr <- forecast::forecast(fit, h=5)
```
`model="ANN"` specifies the simple exponential smoothing model.


```{r}
plot(fr)
```

Look at the estimates

```{r}
fit
```

#### The weighting function {-}

```{r ann.weighting, echo=FALSE}
alpha <- coef(fit)[1]
wts <- alpha*(1-alpha)^(0:23)
plot(1987:1964, wts/sum(wts), lwd=2, ylab="weight", xlab="", type="l")
title("Weighting for simple exp. smooth of anchovy")
```

---

### Decomposing your model fit

Sometimes you would like to see the smoothed level that the model estimated. You can see that with `plot(fit)` or `autoplot(fit)`.

```{r fig.height=4}
autoplot(fit)
```


